---
title: "PDXDataSciReccomender"
author: "Charles Howard"
date: "November 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width=9.5, fig.height=4.5, comment=NA, rows.print=16, out.width = "\\textwidth")
```

## Overview

The goal is to build a recommendation engine for games.  The R package arules is used to mine associations between lists of items.  The arulesViz package has plot methods to visualize relationships between items.

I started with the original set of 834415 rows and 3 columns.  The arules package requires nominal variables be converted to factors and continuous variables to be discretized.  I followed examples given in the following webpage:
http://michael.hahsler.net/research/arules_RUG_2015/demo/

R code follows:

```{r prepping_data dat}
library(arules)
library(arulesViz)
#library(Matrix)  if needed
datdir<-"C:/Users/Charles/Documents/PDXDataSciRecommender/"
setwd(datdir)
dat<-read.csv(paste(datdir,"boardgame-ratings.csv",sep=""))
# sorting data by 1.) UserId, then 2.) gameID
dat<-dat[order(dat$UserID,dat$gameID),]
# determining groupings by UserID
usergrping<-grouping(dat$UserID)
userid.ends<-attr(usergrping,"ends")
userid.starts<-c(1,userid.ends[1:(length(userid.ends)-1)]+1)
userid.counts<-diff(userid.starts)
# convert to factors
dat[,"UserID"]<-factor(dat[,"UserID"])
dat[,"gameID"]<-factor(dat[,"gameID"])
# discretize ratings
dat[,"rating"]<-discretize(dat$rating,method="interval",categories=5)
# for first attempt, I create a list of gameID's by UserID
translist<-lapply(1:length(userid.ends),function(n){
  rws<-userid.starts[n]:userid.ends[n]
  x<-dat$gameID[rws]
})
# the transaction class is the primary one used for arules
datrans<-as(translist,"transactions")
```

Each list in translist is a "transaction".  In this instance, a list of gameID's for UserID "1".
```{r}
print(translist[[1]])
```
Summary of the datrans transactions object.

```{r}
summary(datrans)
```

Some standard measures for item lists are support and confidence.  Support is the proportion of a given item list in the data.  Confidence is a conditional probability type measure.  The confidence of item set A => item set B is:
  support(item set A) <union> support(item set B)/support(item set A)  
I arbitrarily chose a target of 1000 to arrive at a support value.

```{r trans_support}
# find a support level
sup<-1000/nrow(datrans)
print(paste("support is ",sup,sep=""))
itemsets <- apriori(datrans, parameter = list(target = "frequent",
                                            supp=sup, minlen = 3))
inspect(head(sort(itemsets), n=10))
```

There's a cool graph method.

```{r support_graph}
plot(head(sort(itemsets, by = "support"), n=10), method = "graph", control=list(cex=.8))

```

The first grouping in the table above.

```{r support_graph2}
plot(head(sort(itemsets, by = "support"), n=1), method = "graph", control=list(cex=.8))

```
The first and second groupings in the table above.

```{r support_graph3}
plot(head(sort(itemsets, by = "support"), n=2), method = "graph", control=list(cex=.8))

```
...and so on...
```{r support_graph4}
plot(head(sort(itemsets, by = "support"), n=3), method = "graph", control=list(cex=.8))

```
